Enabling Ambient Backscatter Using Low-cost Software-defined-radio

### 2017-04-03 paper.tex
Wrote a lot about the receiver. Done for today.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

@@

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

@@

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

@@

\subsection{RTL2832U}

@@

\subsection{Receiver}
\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{fig/receiver_arch}
\caption{This figure shows the receiver architecture from an abstract point of view.}
\label{fig_receiver_arch}
\end{figure}

@@

>>>>

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

@@

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

@@

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

@@

\subsection{RTL2832U}
\label{sub:rtl2832}

@@

\subsection{Receiver}
\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{fig/receiver_arch}
\caption{This figure shows the receiver architecture from an abstract point of view.}
\label{fig:receiver_arch}
\end{figure}
The receiver was the most work intensive part of the design. C++ has been used to implement the receiver. We provide the code publicly available under \url{https://github.com/s3xm3x/backscatterBASKReceiver}. In figure \ref{fig:receiver_arch} one can see the architecture of the receiver from an abstarct point of view. A highly sophisticated bus system has been developed to exchange data between the different components of the system. The signal flow is from the left to the right. \textit{Librtlsdr} (cf. \cite{github:librtlsdr}), which is based on \textit{libusb}, is used to transfer the data from the TV stick into our program. As described under \ref{sub:rtl2832}, the RTL2832 mixes down the high frequency to a intermediate frequency (both values are in SW adjustable). Subsequently the data is mixed with sine and cosine, low-pass filtered and finally transfered in the digital world. This data is available as two 8 Bit values. It is a well-known fact, that this technique, which is commonly known as I/Q demodulation, results in a real and an imagenery part. With this two values phase and absolute value of the data can be determined for every sample. So every sample consists out of a real and and imaginary part. And can be seen as:
\begin{equation}
	I+j\,Q = abs(I,Q) \cdot e^{ang(I,Q)} 
\end{equation} 
So the first block, which is entitled with RTL2832U, provides the interface to the TV dongle. This block is represented as one object. It can set the frequency \ensuremath{f_{tuned}}, where the receiver is listening. Also it can set the sampling frequency \ensuremath{f_{samp}} and the analog gain of the tuner. The raw samples are pushed to the bus. 
Demodulator is the block called, which is responsible to convert the sampled signals into something, which looks like ones and zeroes. Thereforre the real and imgainary values first have to be deinterleaved out of the sample message. Now they are represented as integer values between 0 and 255 (8 bit). So one converts them to floats, whereas 127.5 equals 0. After that the values have to be downsampled. Usually we were sampling with 250 kHz. This is still much more than we need assuming an additive white gausian noise (AWGN) channel with a \ensuremath{S/N = 10\,dB} (cf. equation \ref{eq:awgn}). So we decided to reduce the sampling frequency to 25 kHz.
\begin{equation}
	\label{eq:awgn}
	C=B \cdot log_2(1+\frac{S}{N})= 25\,kHz \cdot log_2(11) \approx 86.5\,kbit/s
\end{equation}     
To do this it is important to use an anti-aliasing filter, as the Shannon-Nyquist theorem has to be satisfied: 
\begin{equation}
	\label{eq:nyquist}
	f_{sampling} > 2 \cdot f_{highestfregencyinsignal}
\end{equation}
A FIR filter naivly implemented with floats has been used to achieve this. 10 kHz is the cut-off frequency of the first filter. All filters have been constructed with the internal filter design tool of GNU radio. After downsampling another filter is used to suppress noise. This filter has a cut-off frequency of 3 kHz. The last demodulation step is to take the signal, rectify it (cf. equation \ref{eq:rectify}) and decide with a software-defined Schmitt trigger whether a sample is a 1 or a 0. 
\begin{equation}
	\label{eq:rectify}
	abs(I,Q) = \sqrt{I^2 +  Q^2}
\end{equation} 
These samples are broadcasted on the bus. A registered listener receives the processed samples. This registered listener is the decoder, which is represented as an object as well. Thereupon the decoder decides when a frame is sent or when the channel is idle. On the basis of a defined threshold the decoder decides wether a certain sample pattern of ones and zeros is a 1 bit or a 0 bit. The baudrate value and the sampling rate value lead to a certain length of one bit. For example a sample rate \ensuremath{f_{sampling}} of 25 kHz combinded with a baudrate of 1 kbit/s lead to 25 samples/bit.

@@

\bibitem{github:librtlsdr}
Steve Markgraf et. al., \emph{librtlsdr}, \url{https://github.com/steve-m/librtlsdr}. Github: 2012-2017.
